---
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r global_options, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning =FALSE, echo= FALSE)

```

Sao Paulo food distribution by Dlaivison R. Silva
========================================================
> The Objective of this exploration is to identify Sao Paulo market for foods and distribution
based on the information provided by openstreetmap.org

```{r echo=FALSE, message=FALSE, warning=FALSE, packages}



library(ggplot2)
library(ggmap)
library(maptools)
library(maps)
library(stringr)
library(dplyr)
library(gridExtra)
library(jsonlite)
library(OpenStreetMap)
library(rgdal)

```

```{r echo=FALSE, message=FALSE, warning=FALSE, Load_the_Data}
# Load the Data

dados <-fromJSON("saopaulo.osm.json", flatten=TRUE)
colnames(dados)



#Filtering the 1st stage to only keep food places

food_places <- subset(dados,(dados$Business=="restaurant" | dados$Business=="cafe" | dados$Business=="fast_food"))



# Local distribuition of food places

 ggplot(data=food_places, aes(food_places$Business)) + geom_histogram(stat='count',colour='black', fill='blue') + xlab("Number of food places in Sao Paulo central area")
 

# location of the food places in Sao Paulo Map

location_food_places <- data.frame(food_places$Name,food_places$Business,food_places$latitude,food_places$Longitude)

xy <- dataframe(location_food_places$latitude,location_food_places$longitude)
map <- openmap(c(-23.4518,-46.8195),c(-23.6718,-46.4260))
plot(map,raster=TRUE)


# Restaurant concentration per street in Sao Paulo


ggplot(data=food_places, aes(food_places$address.street)) + geom_histogram(stat='count',colour='black', fill='blue') + xlab("Concentration per street") + coord_flip()
 


 
  





```


# Univariate Plots Section



```{r echo=FALSE, message=FALSE, warning=FALSE, Univariate_plots_section}

str(chocolate)
summary(chocolate)
cho<-chocolate

```

# Cocoa rating distribution scale

```{r echo=FALSE, message=FALSE, warning=FALSE}

plot(aes(Rating),data=cho) + geom_histogram(color="black", fill='yellow') +
  xlab("Rating levels")

```

> This rating resides besides the lowest value =0 to highest value=5


# Percentage levels of chocolate sold in the world

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(as.numeric(substr(as.character(Cocoa.Percent),0,2))/100), data=cho)+ geom_histogram(stat='count',colour='black', fill='blue') + scale_x_continuous(breaks=seq(0.1,1,0.1),labels=scales::percent) + 
xlab('Percentage levels of cocoa') 

```

> Above the percentage levels of cocoa in the manufacturing.It shows that the 
concentration of 70% is by far the more produced in the world. 
This is because the 70% of cocoa is also used in other areas such pastry 
and ice cream industry

# Main chocolate company locations

```{r echo=FALSE, message=FALSE, warning=FALSE}

data(wrld_simpl)
myCountries = wrld_simpl@data$NAME %in% chocolate$Company.Location
plot(wrld_simpl, col = c(gray(.80), "blue")[myCountries+1]) + 
  title ("Main Chocolate Company locations")

```

>The map above shows the location of the main chocolate manufactures in the world.


# Main Cocoa bean exporters

```{r echo=FALSE, message=FALSE, warning=FALSE}

data(wrld_simpl)
myCountries = wrld_simpl@data$NAME %in% chocolate$Broad.Bean.Origin
plot(wrld_simpl, col = c(gray(.80), "red")[myCountries+1])+
  title("Main cocoa bean exporters")

```

> As showed above the main cocoa producers are countries located closer to Ecuador where the cocoa developes better


# The top Manufacturing countries with the best rating chocolate  (rating >4)

```{r echo=FALSE, message=FALSE, warning=FALSE}

choc_top_best <- subset(cho,cho$Rating>=4) %>% group_by(Company.Location) %>% summarise(count=n())

ggplot(aes(x=reorder(Company.Location,count), y=count),data=choc_top_best) +
  geom_bar(stat='identity',fill='blue') +
  coord_flip() + xlab('Company Locations')

```

>  This plot shows that considering the most of all top rated chocolate (above 4) have their locations
in USA , France, Italy, Canada, Belgium and Switzerland. There are more american chocolate companies that produces
a very good rated chocolated than other countries. 

# The top  Manufacturing countries with the lower rating Chocolate

```{r echo=FALSE, message=FALSE, warning=FALSE}

choc_top_lower <- subset(cho,cho$Rating<=1.5) %>% group_by(Company.Location) %>% summarise(count=n())
ggplot(aes(x=reorder(Company.Location,count), y=count),data=choc_top_lower) +
  geom_bar(stat='identity',fill='blue') +
  coord_flip() + xlab('Company Locations')
```

> This plot shows that for companies with lower rating level, there are also USA, France and Belgium companies.
But there is something here that must be also considered. Sometimes the type of cocoa produced is not at high quality
because its destination is to be used not as decoration, eg: chocolate sprinkles. So it could explain why some 
countries with high qualified rates appears also here.


# Bean Type distribution 

```{r echo=FALSE, message=FALSE, warning=FALSE}

ggplot(aes(cho$Bean.Type),data=cho) + geom_histogram(stat='count', fill='red')+  
  coord_flip() + xlab('Bean Type')

```

> This plot shows the Bean type sold in the world


# Main manufacturing companies

```{r echo=FALSE, message=FALSE, warning=FALSE}

choc_companies <- cho %>% group_by(CompanyÂ...Maker.if.known.) %>% 
  summarise(count=n())

ggplot(aes(reorder(CompanyÂ...Maker.if.known.,count), y=count),data=subset(choc_companies, 
                                                            count>10)) + 
  geom_bar(stat='identity', fill='darkgreen')+  
  coord_flip() + xlab('Company name')

```

> In the plot above SOMA (Canadian company) is the main producer of the samples collected 
in this research.



# Univariate Analysis



### What is the structure of your dataset?
> The dataset contains data about the main chocolate producers, main cocoa bean exporters and the quality of beans/chocolate classified in an international rate.



### What is/are the main feature(s) of interest in your dataset?

> To understand and classify the best chocolate beans and relate them to the main manufacturers. The main question to answer here is "Does the best bean means the best chocolate" ? 


### What other features in the dataset do you think will help support your \investigation into your feature(s) of interest?

> Based on the content of the dataset I also will determine if some companies still produces the best rating chocolates in the world or if there will be new companies in the field.

### Did you create any new variables from existing variables in the dataset?
> yes. I've created a varible called Bean.filter to adjust some repeatable cocoa beans types.


### Of the features you investigated, were there any unusual distributions? 
> So far the distributions seems to be according what I expected.



# Bivariate Plots Section


```{r echo=FALSE, message=FALSE, warning=FALSE, Bivariate_Plots}

```

# Percentage chocolate levels x Rating
> In order to perform such analysis I've created another variable called Bean.Filter where I have filtered the values from the Beans just to specify one specie.


```{r echo=FALSE, message=FALSE, warning=FALSE}

#1st filter
cho$Bean.filter <-substr(cho$Bean.Type,0,str_locate(cho$Bean.Type,",")[,1]-1)
cho$Bean.filter <-ifelse(is.na(cho$Bean.filter)
                         ,as.character(cho$Bean.Type),cho$Bean.filter)

#2nd  filter
cho$Bean.temp <-substr(cho$Bean.filter,0,
                       str_locate(cho$Bean.filter,"\\(")[,1]-1)
cho$Bean.temp <-ifelse(is.na(cho$Bean.temp),
                       as.character(cho$Bean.filter),cho$Bean.temp)
cho$Bean.filter<-cho$Bean.temp
cho$Bean.filter<-trimws(cho$Bean.filter)
cho$Bean.temp<-NULL




ggplot(aes(x=as.numeric(substr(as.character(Cocoa.Percent),0,2))/100,
           y=Rating), data=cho)  + 
  geom_jitter(alpha = 1, size = 1.5, position = 'jitter') + 
  scale_x_continuous(breaks=seq(0.1,1,0.1),labels=scales::percent)+ 
  xlab('Percentage levels of cocoa') + 
  theme_light()+ labs(color="Bean Types") +
  labs(title="Percentage chocolate levels x Rating")

```

> This chart shows the percentage levels of cocoa vs the Rating. There is a concentration of points
around the percentage of 70%. This means the percentage of 70% is the most important and qualified for 
chocolate processing






# The best Manufacturing countries based on the mean Rating 

```{r echo=FALSE, message=FALSE, warning=FALSE}

choc_rates <- cho %>% group_by(Company.Location) %>% 
  summarise(rate=mean(Rating))
choc_rates <-choc_rates[order(-choc_rates$rate),]

ggplot(aes(x=reorder(Company.Location,rate), y=rate), data=head(choc_rates,10))+ geom_bar(stat='identity',fill='darkblue') +coord_flip() +
  xlab('Company Location') + 
  labs(title="The best Manufacturing countries based on the mean Rating")

```

> Plotting the mean value we can find an interesting fact. Chile although not participating in the top 5 best level rating, at mean values is the country that produces the best rating values.


# Top 10 cocoa best bean producers (non Manufactures). Producers with the best cocoa beans

```{r echo=FALSE, message=FALSE, warning=FALSE}

choc_prod <- cho %>% group_by(Broad.Bean.Origin) %>% summarise(count=n())
choc_prod <-choc_prod[order(-choc_prod$count),]

ggplot(aes(x=reorder(Broad.Bean.Origin,count), y=count), data=head(choc_prod,10)) + geom_bar(stat='identity',fill='darkred') +coord_flip() +xlab('Bean Origin')

```

> I've checked this info twice and seems that the best cocoa beans comes really from Venezuela. I know also that Jacarta's Island (Indonesia) produces a very good cocoa and is not mentioned in this report, but all other countries are really considered the best cocoa bean producers.


# The worst Manufacturing countries based on mean Rating

```{r echo=FALSE, message=FALSE, warning=FALSE}

choc_rates <-choc_rates[order(choc_rates$rate),]
ggplot(aes(x=reorder(Company.Location,-rate), y=rate), 
       data=head(choc_rates,10)) + 
  geom_bar(stat='identity',fill='darkblue') +coord_flip() +
  xlab('Company Location')

```

> This plot shows countries where the mean rate is considered low


# The best bean types based on the mean rating values

```{r echo=FALSE, message=FALSE, warning=FALSE}

choc_bean <- subset(cho,cho$Bean.filter!="") %>% group_by(Bean.filter) %>% 
  summarise(total=mean(Rating))
ggplot(aes(x=reorder(Bean.filter,total),y=total),data=choc_bean) +
  geom_bar(stat='identity',fill='red') + coord_flip() + 
  xlab('Bean Type')

```

> According to the plot the best bean types based on mean rating values are "Blend-Forasteiro" and "Amazon mix"


# Percentage levels x Rating

> In order to perceive better some relations , the category of Cocoa Percent will be update to the following 
values
.Less_than_70    (x < 60)
.Around_70       (60<=x<=80)
.More_than_70    (x > 80 )

```{r echo=FALSE, message=FALSE, warning=FALSE}

cocoa_percent = as.numeric(substr(cho$Cocoa.Percent,0,2))

cho$levels <- ifelse(cocoa_percent>80,"More_than_70", 
                     ifelse((cocoa_percent>=60 & cocoa_percent<=80),
                            "Around_70","less_than_70")) 

cho$levels <- factor(cho$levels,ordered=TRUE,levels=c("less_than_70","Around_70","More_than_70"))

cho$level  <- ifelse(cocoa_percent>80, 100,
                     ifelse((cocoa_percent>=60 & cocoa_percent<=80),
                            70,30)) 



ggplot(aes(x=cho$Rating,y=cho$level, fill='blue'),data=cho)+ geom_jitter(alpha=0.4,
                                                            fill='blue')+
  geom_smooth(fill='green') + xlab('Rating') + ylab('Cocoa percentage') 


cor.test(as.numeric(cho$level==30),cho$Rating)
cor.test(as.numeric(cho$level==70),cho$Rating)
cor.test(as.numeric(cho$level==100),cho$Rating)



```

> For values around 70% the correlation has a value closer to 0.2.  



# Main manufacturing companies based on Rating

```{r echo=FALSE, message=FALSE, warning=FALSE}


ggplot(aes(reorder(CompanyÂ...Maker.if.known.,Rating), y=Rating), 
       data=subset(cho,Rating>=4))+
  geom_bar(stat='identity', fill='red')+  
  coord_flip() + xlab('Company name')  


```

> Based in the samples collected, company SOMA has the majority of chocolate rated
better. It could be related to the quantity of samples collected.




# Main manufacturing companies based on Mean Rating

```{r echo=FALSE, message=FALSE, warning=FALSE}

choc_rating <- cho %>% group_by(CompanyÂ...Maker.if.known.) %>% 
  summarise(Total=mean(Rating))

ggplot(aes(reorder(CompanyÂ...Maker.if.known.,Total),y=Total),
       data=subset(choc_rating, Total>3.6)) + 
  geom_bar(stat='identity', fill='red')+  
  coord_flip() + xlab('Company name')

```

> Considering the mean value, the company Tobago Estate has the better ratings.




# Bivariate Analysis


### Talk about some of the relationships you observed in this part of the \
investigation. How did the feature(s) of interest vary with other features in \
the dataset?
> I could identify the best ratings seems to be related to the percentage of Cocoa. For example the best ratings are around 65-75% of chocolate percentage.

### Did you observe any interesting relationships between the other features \
(not the main feature(s) of interest)?
> I noticed that some companies have more samples analised on this research that could contribute to have
a lack of data interpretation in terms of quality.



### What was the strongest relationship you found?
> Percentage of cocoa 65-75% indicates a best cocoa (chocolate)



# Multivariate Plots Section


# Percentage chocolate levels x Rating
> In order to perform such analysis I've created another variable called Bean.Filter where I have filtered the values from the Beans just to specify one specie.


```{r echo=FALSE, message=FALSE, warning=FALSE}

#1st filter
cho$Bean.filter <-substr(cho$Bean.Type,0,str_locate(cho$Bean.Type,",")[,1]-1)
cho$Bean.filter <-ifelse(is.na(cho$Bean.filter)
                         ,as.character(cho$Bean.Type),cho$Bean.filter)

#2nd  filter
cho$Bean.temp <-substr(cho$Bean.filter,0,
                       str_locate(cho$Bean.filter,"\\(")[,1]-1)
cho$Bean.temp <-ifelse(is.na(cho$Bean.temp),
                       as.character(cho$Bean.filter),cho$Bean.temp)
cho$Bean.filter<-cho$Bean.temp
cho$Bean.filter<-trimws(cho$Bean.filter)
cho$Bean.temp<-NULL




ggplot(aes(x=as.numeric(substr(as.character(Cocoa.Percent),0,2))/100,
           y=Rating,color=Bean.filter), data=cho)  + 
  geom_jitter(alpha = 1, size = 1.5, position = 'jitter', 
              aes(colour=cho$Bean.filter)) + 
  scale_x_continuous(breaks=seq(0.1,1,0.1),labels=scales::percent)+ 
  xlab('Percentage levels of Cocoa') + 
  theme_dark()+ labs(color="Bean Types")+
  labs(title="Percentage levels of cocoa x Rating and Bean Type")





```

> This chart shows the percentage levels of cocoa vs the Rating. There is a concentration of points
around the percentage of 70%. This means the percentage of 70% is the most important and qualified for 
chocolate processing


# Countries and respective rating levels

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=9, pair_plots}


ggplot(aes(y=chocolate$Company.Location,x=Rating), data=cho)  +
  geom_jitter(alpha = 1, size = 1, position = 'jitter', 
                aes(colour=cho$levels)) + scale_x_continuous() +
  xlab('Rating levels of cocoa per Country') +ylab('Company Location')+
labs(color='Percentage Levels', title='Countries x Rating Levels') 


```

> The chart shows there are a high concentration of companies more in USA with also rating levels between 2.5 -4. And peak levels of 5 for countries such Italy and Japan
The traditional chocolate and famous manufactures such as Switzerland and Belgium contains rating between 1-4.5

> The chart above shows the best rating are for chocolate percentage between 0.65-0.80



# Correlation factor between percentage of cocoa and cocoa rating

```{r echo=FALSE, message=FALSE, warning=FALSE}



ggplot(aes(x=as.numeric(substr(as.character(Cocoa.Percent),0,2))/100,y=Rating,
           color=Bean.filter), data=cho)  + 
  geom_jitter(alpha = 1, size = 1.5, position = 'jitter', 
              aes(colour=cho$Bean.filter)) + 
  scale_x_continuous(breaks=seq(0.1,1,0.1),labels=scales::percent) +
  xlab('Percentage levels of cocoa') + 
  theme_dark()+ labs(color="Bean Types") +geom_smooth(method='lm')+
  labs(title="Correlation factor between percentage of cocoa and cocoa rating")

cor.test(as.numeric(substr(as.character(cho$Cocoa.Percent),0,2)),cho$Rating)

```


# Combining Mean Rating, Rating, cocoa percent and Company 
>Items to consider:
> Rates > 3
> Percentage between 65-75



```{r echo=FALSE, message=FALSE, warning=FALSE, fig.width=15, fig.height=15, choc_rates}


choc_rating <- cho %>% group_by(CompanyÂ...Maker.if.known.) %>% 
  summarise(Total=mean(Rating),count=n(),
            Percent=mean(as.numeric(substr(Cocoa.Percent,0,2))))

choc_rating <- subset(choc_rating, choc_rating$Total>3.5)
choc_rating <- subset(choc_rating, (choc_rating$Percent>=65))
choc_rating <- subset(choc_rating, (choc_rating$Percent<=85))                      
                      



p1 <- ggplot(aes(reorder(CompanyÂ...Maker.if.known.,Total),y=Total),
       data=choc_rating) + 
  geom_point(alpha=1, aes(color=choc_rating$Percent))+  
  coord_flip() + xlab('Company name') + 
  labs(color='Cocoa Percent',title="Analysis per Company/Rating") +
  theme_light() +ylab('Cocoa Rating')

p2<-  ggplot(aes(reorder(CompanyÂ...Maker.if.known.,Total),y=count),
       data=choc_rating) + 
  geom_point(alpha=1, aes(color=choc_rating$Percent))+  
  coord_flip() + xlab('Company name') + 
  labs(color='Cocoa Percent', title="Analysis per Company/Counting") +
  theme_light() +ylab('Count of Chocolate')


grid.arrange(p1,p2)

```


> From the plot above we can see that the company Tobago Estate has comparing to Soma company less samples, but
these samples has higher rating. So Soma has appearead before also because it has more quantities of samples comparing to Tobago Estate.


# Multivariate Analysis

### Talk about some of the relationships you observed in this part of the \
investigation. Were there features that strengthened each other in terms of \
looking at your feature(s) of interest?

> We can perceive after investigating all the plots that when the chocolate has the percentage levels closer to 70% , the rating level are closer to the best. Probably this percentage is the preferred for the best brands for processing.
The 70% chocolate is a pure chocolate and cocoa butter.


### Were there any interesting or surprising interactions between features?

> The surprise was to see countries like Belgium with lower rates. Later I noticed that such lower rates probably is related to the type of cocoa and its type of use. Probably is a low quality cocoa destinated to be used to produce chocolate derivatives such as chocolate sprinkles and liquor.

### OPTIONAL: Did you create any models with your dataset? Discuss the strengths and limitations of your model.\

> No. I don't. There were no such many numeric variables that I could related. The only numeric and important variables were the percentage of cocoa and rating. The most of all data are qualitative and not quantitative.
And the unique variable that has an interval of linear relation is the cocoa percent x rating (at the interval of 65-85%)

------

# Final Plots and Summary

> **Tip**: You've done a lot of exploration and have built up an understanding
of the structure of and relationships between the variables in your dataset.
Here, you will select three plots from all of your previous exploration to
present here as a summary of some of your most interesting findings. Make sure
that you have refined your selected plots for good titling, axis labels (with
units), and good aesthetic choices (e.g. color, transparency). After each plot,
make sure you justify why you chose each plot by describing what it shows.

### Plot One
```{r echo=FALSE, message=FALSE, warning=FALSE,Plot_One}


ggplot(aes(x=as.numeric(substr(as.character(Cocoa.Percent),0,2))/100,
             y=Rating,color=Bean.filter), data=cho)  + 
  geom_jitter(alpha = 1, size = 1.5, position = 'jitter',
                aes(colour=cho$Bean.filter)) +
  scale_x_continuous(breaks=seq(0.1,1,0.1),labels=scales::percent) + 
  xlab('Cocoa Rating') + 
  theme_dark()+ labs(color="Bean Types") +geom_smooth(method='lm')+
  labs(title="Percentage Leves vs Rating and Bean Type")


```

### Description One
> This  plot shows the correlation factor between the rating and percentage of Cocoa. For low rate levels there is linear correlation between these values. But at range of 65-75% the correlation is closer and the rating levels are better.


### Plot Two
```{r echo=FALSE, message=FALSE, warning=FALSE,Plot_two}



ggplot(aes(x=as.numeric(substr(as.character(Cocoa.Percent),0,2))/100,
           y=Rating), data=cho)  + 
  geom_jitter(alpha = 1, size = 1.5, position = 'jitter') + 
  scale_x_continuous(breaks=seq(0.1,1,0.1),labels=scales::percent)+ 
  xlab('Percentage levels of cocoa') + 
  theme_light()+ labs(color="Bean Types") +
  labs(title="Percentage chocolate levels x Rating")

```

### Description Two
> This plot shows that the best ratings and samples collected are around 70%. This results shows that most of all
companies prefers to have chocolates with cocoa percent around 70%

### Plot Three
```{r echo=FALSE, message=FALSE, warning=FALSE,Plot_Three}

ggplot(aes(x=reorder(Company.Location,rate), y=rate), data=head(choc_rates,10))+ geom_bar(stat='identity',fill='darkblue') +coord_flip() +
  xlab('Company Location') + 
  labs(title="The best Manufacturing countries based on the mean Rating")

```

### Description Three
> This plot shows a summary of the manufactures that produces the best chocolate in average. My surprise is to see Chile in the top, although the chart represents the mean value, not the top value. 

------

# Reflection

> During my research several types of analysis have been performed. From univariate analysis the focus was to see the distribution of the main variables (rating, percent of cocoa, number of samples) per manufacturing company, country and Bean exporter. At this level some important points were take in consideration for the second level of analysis at Bivariate section.
  For Bivariate analysis, the main focus was to find a correlation between certain values. The variables Cocoa Percent and Rating has shown  some linearity between the interval of 65-75 % (cocoa percent).  The bean types Amazon and Blend Forastero were the best types found on this research
At Multivariate analysis the main target was to determine the correlation factor between the cocoa percent and rating and have the linear function to interpret the values. No model has been built due the limitation of variables and data.
   There were several challenges on this project. As the data set has no numerical variables I needed to built some intermediate variables to have a better interpretation of the data. These intermediate variables were used to evaluate and better determine the correlation and judge the best companies, countries and bean types.
   I learned also how to manipulate the geographic maps in my research. Because plotting only the countries's names was not enough than plotting the regions. In terms of data analysis the data interpretation is a key.
   I believe this dataset must be improved and should contains new variables, such as bean granularity, type_of_use (for which use the chocolate is designated),  acid levels , etc. In my research I have noticed the companies focus more in percentage level of 70% than other levels, so therefore this percentage should be considered the main in terms of research.
   